Whitepaper: Contract Modules

### 2021-01-19 meeting-notes/20201108-expose-input-language.tex
Merge Master

\begin{center}
  \begin{tabular}{|l|l|}
    Michelson & OCaml \\\hline
    \texttt{string}& \texttt{string}  \\
    \texttt{nat} & ? (\texttt{int}) \\
    \texttt{int} & \texttt{int} \\ 
    \texttt{bytes}& ? \\
    \texttt{bool}&\texttt{bool} \\
    \texttt{unit}&\texttt{unit} \\
    \texttt{list t}&\texttt{t list} \\
    \texttt{pair l r}&\texttt{l * r} \\
    \texttt{option t}     &\texttt{t option} \\
    \texttt{or l r}&\texttt{[`Left of l | `Right of r]} \\
    \texttt{set t}&? \texttt{Set.Make (T).t} \\
    \texttt{map k t}&? \texttt{`t Map.Make(K).t} \\
    \texttt{big\_map k t}&? \texttt{`t Map.Make(K).t} \\
    \hline
  \end{tabular}
\end{center}

@@

We can represent Michelson parameter values with the following OCaml type using GADTs.
\begin{lstlisting}[language=Caml]
type _ Mvalue =
  | Mstring : string -> string Mvalue
  | Mint  : int -> int Mvalue
  | Mbool : bool -> bool Mvalue
  | Munit : unit -> unit Mvalue
  | Mnil  : 't list Mvalue
  | Mcons : 't Mvalue * 't list Mvalue -> 't list Mvalue
  | Mpair : 'l Mvalue * 'r Mvalue -> ('l * 'r) Mvalue
  | Mnone : 't option Mvalue
  | Msome : 't Mvalue -> 't option Mvalue
  | Mleft : 'l Mvalue -> [`Left of 'l | `Right 'r] Mvalue
  | Mright : 'r Mvalue -> [`Left of 'l | `Right 'r] Mvalue
\end{lstlisting}

\begin{lstlisting}[language=Caml]
type _ Mtype = 
  | Tstring : string Mtype
  | Tint    : int Mtype
  | Tbool   : bool Mtype
  | Tunit   : unit Mtype
  | Tlist   : 't Mtype -> 't list Mtype
  | Tpair   : 'l Mtype * 'r Mtype -> ('l * 'r) Mtype
  | Toption : 't Mtype -> 't option Mtype
  | Tor     : 'l Mtype * 'r Mtype 
            -> [`Left of 'l | `Right of 'r] Mtype
\end{lstlisting}
Instead of using polymorphic variants, one could also define a binary
sum type like Haskell's Either type.
\begin{lstlisting}[language=Caml]
type ('a, 'b) either = Left of 'a | Right of 'b
\end{lstlisting}

@@

Or given a contract for index type \lstinline/t/, we'd have an
invocation function
\begin{lstlisting}[language=Caml]
invoke : 't Mtype Tezos.contract -> 't -> result Tezos.apimonad
\end{lstlisting}
with an internal serialization function
\begin{lstlisting}[language=Caml]
serialize : 't Mtype -> 't -> string
\end{lstlisting}
Not sure if deserialization is also needed, but it might be necessary
to decode the current state of a contract. Deserialization may fail
and thus raise an exception.
\begin{lstlisting}[language=Caml]
deserialize : 't Mtype -> string -> 't
\end{lstlisting}

@@

Query the Tezos blockchain for the type of a contract and its
entrypoints.
\begin{lstlisting}[language=Caml]
make_typed_contract : contract_hash -> 't Mtype -> 't Contract
\end{lstlisting}
The function call \lstinline/make_typed_contract hash mty/ would query
the node (via client lib or rpc) for the type of the contract
\lstinline/hash/ and check it against \lstinline/mty/.
This call fails if the type \lstinline/mty/ does not match the
stored type of the contract.
The returned value contains the serializer from \lstinline/'t -> string/.
\begin{lstlisting}[language=Caml]
invoke_typed : 't Contract -> 't -> result Tezos.apimonad
\end{lstlisting}
\end{document}

>>>>

\begin{center}
  \begin{tabular}{|l|l|l}
    Michelson type & OCaml type & Comment \\\hline
    \texttt{string}& \texttt{string}  \\
    \texttt{nat} & ? (\texttt{Z.t}) & \texttt{Zarith} \\
    \texttt{int} & \texttt{Z.t} & \texttt{Zarith}\\ 
    \texttt{bytes}& \texttt{bytes} \\
    \texttt{bool}&\texttt{bool} \\
    \texttt{unit}&\texttt{unit} \\
    \texttt{list t}&\texttt{t list} \\
    \texttt{pair l r}&\texttt{l * r} \\
    \texttt{option t}     &\texttt{t option} \\
    \texttt{or l r}&\texttt{[`Left of l | `Right of r]} \\
    \texttt{set t}&? \texttt{Set.Make (T).t} \\
    \texttt{map k t}&? \texttt{T Map.Make(K).t} \\
    \texttt{big\_map k t}&? \texttt{T Map.Make(K).t} \\
    \texttt{mutez} & \texttt{int64} \\
    \hline
  \end{tabular}
\end{center}

@@

We can represent Michelson parameter values with the following OCaml GADT.
\begin{lstlisting}[language=Caml]
type _ Mvalue =
  | Mstring : string -> string Mvalue
  | Mint  : int -> int Mvalue
  | Mbool : bool -> bool Mvalue
  | Munit : unit -> unit Mvalue
  | Mnil  : 't list Mvalue
  | Mcons : 't Mvalue * 't list Mvalue -> 't list Mvalue
  | Mpair : 'l Mvalue * 'r Mvalue -> ('l * 'r) Mvalue
  | Mnone : 't option Mvalue
  | Msome : 't Mvalue -> 't option Mvalue
  | Mleft : 'l Mvalue -> [`Left of 'l | `Right 'r] Mvalue
  | Mright : 'r Mvalue -> [`Left of 'l | `Right 'r] Mvalue
\end{lstlisting}

\begin{lstlisting}[language=Caml]
type _ Mtype = 
  | Tstring : string Mtype
  | Tint    : int Mtype
  | Tbool   : bool Mtype
  | Tunit   : unit Mtype
  | Tlist   : 't Mtype -> 't list Mtype
  | Tpair   : 'l Mtype * 'r Mtype -> ('l * 'r) Mtype
  | Toption : 't Mtype -> 't option Mtype
  | Tor     : 'l Mtype * 'r Mtype 
            -> [`Left of 'l | `Right of 'r] Mtype
\end{lstlisting}
Instead of using polymorphic variants for expressing alternatives, one could also define a binary
sum type like Haskell's Either type.
\begin{lstlisting}[language=Caml]
type ('a, 'b) either = Left of 'a | Right of 'b
\end{lstlisting}

@@

Or given a contract for index type \lstinline/t/, we'd have an
invocation function
\begin{lstlisting}[language=Caml]
invoke : 't Mtype Tezos.contract -> 't -> result Tezos.apimonad
\end{lstlisting}
with an internal serialization function
\begin{lstlisting}[language=Caml]
serialize : 't Mtype -> 't -> string
\end{lstlisting}
Deserialization is also needed to inspect values in the store of a
contract. Deserialization may fail 
and thus raise an exception.
\begin{lstlisting}[language=Caml]
deserialize : 't Mtype -> string -> 't
\end{lstlisting}

@@

Query the Tezos blockchain for the type of a contract and its
entrypoints.
\begin{lstlisting}[language=Caml]
make_typed_contract : contract_hash -> 't Mtype -> 't Contract
\end{lstlisting}
The function call \lstinline/make_typed_contract hash mty/ would query
the node (via client lib or rpc) for the type of the contract
\lstinline/hash/ and check it against \lstinline/mty/.
This call fails if the type \lstinline/mty/ does not match the
stored type of the contract.
The returned value contains the serializer from \lstinline/'t -> string/. The most
straightforward implementation would be to invoke the untyped sync API
with the serialized parameter.
\begin{lstlisting}[language=Caml]
invoke_typed : 't Contract -> 't -> result Tezos.apimonad
\end{lstlisting}
\end{document}

### 2021-01-19 papers/contract-modules/modules.tex
typo...
Merge Master

Here is a \textbf{contract module} that describes / specifies the above Michelson code.
\begin{lstlisting}[caption={Contract module example},label={lst:contract-module-example}]
contract type MyContract = sig
  paid entrypoint %bid_for_auction (value : int) 
  raises "closed"                 (** auction closed*)
       | "bid too low" as too_low (** bid received is too low *)
       | 42 as the_answer         (** failwith received argument 42 *)
       | (_ : int) as failedwith  (** failwith received an integer arg *)

@@

let close_auction destination ?(fee=50.0) ?(amount=0.0) _ =
  ...
  SyncAPIV0.call_contract
    amount
    destination
    contract_addr
    ~(serialize Tunit)
    fee
\end{lstlisting}
This code example is incomplete and likely incorrect. It contains
fragments from an example implemented on top of the synchronous API.
The call to \lstinline/serialize/ refers to a proposal to improve on
the typing properties of the synchronous API.

>>>>

Here is a \textbf{contract module} that describes / specifies the above Michelson code.
\begin{lstlisting}[caption={Contract module example},label={lst:contract-module-example}]
contract type MyContract = sig
  paid entrypoint %bid_for_auction (v : int) 
  raises "closed"                 (** auction closed*)
       | "bid too low" as too_low (** bid received is too low *)
       | 42 as the_answer         (** failwith received argument 42 *)
       | (_ : int) as failedwith  (** failwith received an integer arg *)

@@

let close_auction destination ?(fee=50.0) ?(amount=0.0) v =
  ...
  SyncAPIV0.call_contract
    amount
    destination
    contract_addr
    ~(serialize Tunit v)
    fee
\end{lstlisting}
This code example is incomplete and likely incorrect. It contains
fragments from an example implemented on top of the synchronous API.
The call to \lstinline/serialize/ refers to a proposal to improve on
the typing properties of the synchronous API.
